Constraint Algebras \\ of Multisymplectic Observables

### 2025-01-19 msconstraints-short-main.tex
Troubleshooting on windows.
Repo simplified due to compilation errors with standalone.

%- HandOut Flag -----------------------------------------------------------------------------------------
	\newif\ifHandout
	%\Handouttrue  %uncomment for the printable version
	%Handling of flags it is not preserved when passing to standalone-subfiles!

%- Document ----------------------------------------------------------------------------------------------
\ifHandout
	\documentclass[handout,10pt]{beamer}   
	\setbeameroption{show notes} %print notes   
\else
	\documentclass[10pt]{beamer}
\fi

%- Packages ----------------------------------------------------------------------------------------------
\usepackage{custom-style}
\usepackage{math}
\usepackage{multicol}
\usepackage{xfrac}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usetikzlibrary{backgrounds}

\renewcommand{\action}{\curvearrowright} 
\makeatletter
\def\blfootnote{\gdef\@thefnmark{}\@footnotetext}
\makeatother

\newcommand{\fgmodule}{\mathfrak{X}_{\g}}
\newcommand{\tgtvanform}{I_{\Omega}(N)}
\newcommand{\tgtvf}{\mathfrak X_N(M)}
\newcommand{\vanvf}{I_{\mathfrak{X}}(N)}%{I_N^{\mathfrak X}}
\newcommand{\Ham}{\mathsf{Ham}}
\renewcommand{\ham}{\text{ham}}

@@

%- Title Page -------------------------------------------------------------------------------------------
\title{ Constraint Algebras \\ of Multisymplectic Observables } 
\subtitle{\href{https://sites.google.com/view/xix-yrw-verona/home?authuser=0}{19th International Young Researchers Workshop on Geometry, Dynamics and Field Theory}}
\author[AMM]{
	\href{https://www.antoniomiti.it/}{Antonio Michele Miti}
	{\small joint work w/ Leonid Ryvkin}
}
\institute[SUR]{
	Sapienza Università di Roma \\
	Rome, Italy 	\\
	\vspace{.5em}
  \begin{tabular}[h]{ccc}
      \href{https://www.mat.uniroma1.it/en}{\includegraphics[width=6cm]{./Logos/Sur_logo}} & & 
      \href{https://civis3i.univ-amu.fr/en/civis3i-alliance-programme}{\includegraphics[width=3cm]{Logos/Civis_logo}}
  \end{tabular}    
}
\date[YGMC] % (optional, should be abbreviation of conference name)
{	
	{\vskip 1ex}
	University of Verona, Italy, January 2025
}

%---------------------------------------------------------------------------------------------------------------------------------------------------
%- Document ----------------------------------------------------------------------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}  % Alternative: \maketitle outside of frame
	\titlepage
	\ifHandout
		\tikz[overlay,remember picture]
		{
	    	%	\node at ($(current page.west)+(1.5,0)$) [rotate=90] {\Huge\textcolor{gray}{\today}};
	    	\node[        
	    		draw,
	    		shape border rotate=90,
			isosceles triangle,
			isosceles triangle apex angle=90,
			fill=yellow]
	        		at ($(current page.north east)-(1,1)$) [rotate=-45] {\textcolor{red}{Handout version}};
		}
	\fi
	\end{frame}
	\addtocounter{framenumber}{-1}
\note{
	%\textbf{\underline{OUTLINE}}:
	%\tableofcontents
	\textbf{Abstract:}
	\\
%
This talk explores key algebraic structures in multisymplectic geometry, a generalization of symplectic geometry where higher-degree $k$-forms replace 2-forms. A central theme is the interplay between symmetries (group actions preserving the differential form) and reduction, which constructs a lower-dimensional space inheriting the geometric structure of the original manifold.

@@

}
%---------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Keywords}
\tikzstyle{every picture}+=[remember picture]
	\begin{columns}
    	\begin{column}{.45\textwidth}
    		\onslide<4->{
			\tikz[baseline]{
		            \node[draw=orange!40,anchor=base,text width=5cm] (s1)
		            {Algebraic structure encoding informations necessary for defining coisotropic reduction in Poisson geometry.};
			}
		}
	\end{column}
    	\begin{column}{.45\textwidth}
    		\onslide<5->{
			 \tikz[baseline]{
		            \node[draw=blue!40,anchor=base, text width=5cm] (s2)
		            {Mathematical formalization of classical measurable quantities.};
			}
		}
	\end{column}
	\end{columns}

@@

\end{frame}
\note[itemize]{
	\item Provide a vague idea of the words that make up the title.
	\item \cite{Dippell2023}
	\item Conventions:
	\\- $M$ and $G$ are connected,
	\\- actions $\theta:G \curvearrowright M$ are always smooth
	\\- $\xi,\eta\in\g$,
	\\- for $\mu\in\Omega^*(M,\g^*)$ and $\xi\in\g$, write
			\[
				\mu_\xi := \langle\mu,\xi\rangle \;{\color{black!50}\in\Omega^*(M)}
			\]
			for the ``$\xi$th component'' of $\mu$.
}
%-------------------------------------------------------------------------------------------------------------------------------------------------

>>>>

%- HandOut Flag -----------------------------------------------------------------------------------------
	\newif\ifHandout
	\Handouttrue  %uncomment for the printable version
	%Handling of flags it is not preserved when passing to standalone-subfiles!

%- D0cum3nt ----------------------------------------------------------------------------------------------
\ifHandout
	\documentclass[handout,10pt]{beamer}   
	\setbeameroption{show notes} %print notes   
\else
	\documentclass[10pt]{beamer}
\fi

%- Packages ----------------------------------------------------------------------------------------------
\usepackage{custom-style}
\usepackage{math}

%- Bibliography (Biber) ----------------------------------------------------------------------------------
\usepackage[backend=biber,style=alphabetic,maxnames=2]{biblatex}
\bibliography{bibfile.bib}

%- L30's ----------------------------------------------------------------------------------------------
\usepackage{quiver} %https://q.uiver.app/ % Not in Miktex!
\usetikzlibrary{nfold}
\usetikzlibrary{decorations.pathmorphing} 
\renewcommand{\Ham}{\mathrm{Ham}}
\newcommand{\Der}{\mathrm{Der}}

@@

%- T1tle P4g3 -------------------------------------------------------------------------------------------
\title{ Constraint Algebras \\ of Multisymplectic Observables } 
\subtitle{\href{https://sites.google.com/view/xix-yrw-verona/home?authuser=0}{19th International Young Researchers Workshop on Geometry, Dynamics and Field Theory}}
\author[AMM]{
	\href{https://www.antoniomiti.it/}{Antonio Michele Miti}
	{\small joint work w/ Leonid Ryvkin}
}
\institute[SUR]{
	Sapienza Università di Roma \\
	Rome, Italy 	\\
	\vspace{.5em}
  \begin{tabular}[h]{ccc}
      \href{https://www.mat.uniroma1.it/en}{\includegraphics[width=6cm]{./Logos/Sur_logo}} & & 
      \href{https://civis3i.univ-amu.fr/en/civis3i-alliance-programme}{\includegraphics[width=3cm]{Logos/Civis_logo}}
  \end{tabular}    
}
\date[YGMC] % (optional, should be abbreviation of conference name)
{	
	{\vskip 1ex}
	University of Verona, Italy, January 2025
}

%--------------------------------------------------------------------------------------------------
%- D0cum3nt ----------------------------------------------------------------------------------------------------------------------------------
\begin{document}
%------------------------------------------------------------------------------------------------
\begin{frame}  % Alternative: \maketitle outside of frame
	\titlepage
	\ifHandout
		\tikz[overlay,remember picture]
		{
	    	%	\node at ($(current page.west)+(1.5,0)$) [rotate=90] {\Huge\textcolor{gray}{\today}};
	    	\node[        
	    		draw,
	    		shape border rotate=90,
			isosceles triangle,
			isosceles triangle apex angle=90,
			fill=yellow]
	        		at ($(current page.north east)-(1,1)$) [rotate=-45] {\textcolor{red}{Handout version}};
		}
	\fi
\end{frame}
\addtocounter{framenumber}{-1}
\note{
	%\textbf{\underline{OUTLINE}}:
	%\tableofcontents
	\textbf{Abstract:}
	\\
%
This talk explores key algebraic structures in multisymplectic geometry, a generalization of symplectic geometry where higher-degree $k$-forms replace 2-forms. A central theme is the interplay between symmetries (group actions preserving the differential form) and reduction, which constructs a lower-dimensional space inheriting the geometric structure of the original manifold.

@@

}
%--------------------------------------------------------------------------------------------------

%------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Keywords}
\tikzstyle{every picture}+=[remember picture]
	\begin{columns}
    	\begin{column}{.45\textwidth}
    		\onslide<4->{
			\tikz[baseline]{
		            \node[draw=orange!40,anchor=base,text width=5cm] (s1)
		            {Algebraic structure encoding informations necessary for defining coisotropic reduction in Poisson geometry.};
			}
		}
	\end{column}
    	\begin{column}{.45\textwidth}
    		\onslide<5->{
			 \tikz[baseline]{
		            \node[draw=blue!40,anchor=base, text width=5cm] (s2)
		            {Mathematical formalization of classical measurable quantities.};
			}
		}
	\end{column}
	\end{columns}

@@

\end{frame}
\note[itemize]{
	\item Provide a vague idea of the words that make up the title.
	\item \cite{Dippell2023}
	\item Conventions:
	\\- $M$ and $G$ are connected,
	\\- actions $\theta:G \curvearrowright M$ are always smooth
	\\- $\xi,\eta\in\g$,
	\\- for $\mu\in\Omega^*(M,\g^*)$ and $\xi\in\g$, write
			\[
				\mu_\xi := \langle\mu,\xi\rangle \;{\color{black!50}\in\Omega^*(M)}
			\]
			for the ``$\xi$th component'' of $\mu$.
}
%------------------------------------------------------------------------------------------------

%------------------------------------------------------------------------------------------------
\section{Symplectic Reduction}
\checkpoint	
%------------------------------------------------------------------------------------------------

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Reduction in Classical Mechanics}

\end{frame}
\note[itemize]{
	\item  In classical mechanics symmetry reduction plays an important role.
	\item Mathematically, this is usually phrased in terms of Marsden-Weinstein reduction
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Geometric Reduction} 
	{Def. Symplectic manifold $(M,\omega)$} $M$ a smooth manifold, $\omega\in \Omega^{2}(M)$ closed and non-degenerate, i.e.:
		\begin{itemize} 	
			\item $d\omega=0$
			\item $\omega^\flat:TM\to T^*M, v\mapsto \iota_v\omega$ is injective
		\end{itemize}

\end{frame}
\note[itemize]{
	\item   Examples:
		\begin{itemize}
			\item orientable 2-manifolds with volume, e.g. $S^2$, $T^2$, $\mathbb R^2$
			\item cotangent bunlde $T^*Q$ of any manifold $Q$, with $\omega=d\theta$. $\theta_\eta(v)=\eta(T\pi_Q(v))$.
			\item coadjoint orbits
		\end{itemize}
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Poisson Reduction} 
{Def. Poisson algebra of observables} Let $(M,\omega)$ be symplectic.
		\begin{itemize}
			\item $X_f:=-(\omega^\flat)^{-1}(df)$  \emph{Hamiltonian vector field} of $f\in C^\infty(M)$.
			\item $\{f,g\}:=\omega(X_f,X_g)=L_{X_f}g$ \emph{Poisson bracket of observables}
		\end{itemize}
	Indeed $(C^{\infty}(M), \cdot, \{-,-\})$ Poisson algebra:
			\begin{itemize}
		\item $\cdot$ is a commutative, associative mutliplication
		\item $\{-,-\}$ is a Lie bracket
		\item $\{f,- \}$ is a derivation
	\end{itemize}\pause
	And $f\mapsto X_f$ is a Lie algebra map. 
\end{frame}
\note[itemize]{
	\item  
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{Marsden-Weinstein-Meyer reduction}
\begin{frame}[fragile]{Marsden-Weinstein-Meyer reduction}
Let $(M,\omega)$ symplectic, $G\circlearrowright M$ be a Lie group action\pause and $\mu:M\to \mathfrak g^*$ such that:\pause
\begin{itemize}
\item  $\mu$ is a momentum, i.e. the infinitesimal action $\xi\mapsto \underline\xi:\mathfrak g\to \mathfrak X(M)$ of $G$ satisfies $X_{\langle\mu,\xi\rangle}=\underline \xi$. \pause
\item $\mu$ is $G$-equivariant, i.e. $\mu(gm)=Ad^*_g\mu(m)$ \pause
\item The action of $G$ on $\mu^{-1}(0)$ is free and proper.\\ (This implies $0$ is reg. val. of $\mu$) \pause
\end{itemize}
Then $\exists!$ symplectic form $\omega_r$ on $M_r=\frac{\mu^{-1}(0)}{G}$ such that $\pi^*\omega_r=i^*\omega$.

% https://q.uiver.app/#q=WzAsMyxbMCwwLCJNIl0sWzMsMCwiXFxtdV57LTF9KDApIl0sWzYsMCwiTV9yIl0sWzEsMCwiaVxcXFwgXFxzdWJzdGFja3tpbmNsdXNpb25+b2Z+c3Vic3BhY2V9IiwyLHsic3R5bGUiOnsidGFpbCI6eyJuYW1lIjoiaG9vayIsInNpZGUiOiJib3R0b20ifX19XSxbMSwyLCJcXHBpXFxcXCBcXHN1YnN0YWNre3F1b3RpZW50fmJ5fnJlbGF0aW9ufSJdXQ==
\[\begin{tikzcd}
	M &&& {\mu^{-1}(0)} &&& {M_r}
	\arrow["\begin{array}{c} i\\ \substack{inclusion~of~subspace} \end{array}"', hook', from=1-4, to=1-1]
	\arrow["\begin{array}{c} \pi\\ \substack{quotient~by~relation} \end{array}", two heads, from=1-4, to=1-7]
\end{tikzcd}\]

\pause
Examples: $\mathbb CP^n$ from $S^1$-action on $\mathbb C^{n+1}$.

\end{frame}

\subsection{Attempting algebraic reduction}

\begin{frame}[fragile]
When the action is not 'free and proper', we can try to mimic this directly using the function algebras:
% https://q.uiver.app/#q=WzAsNixbMCwwLCJNIl0sWzEsMCwiXFxtdV57LTF9KDApPVMiXSxbMywwLCJNX3IiXSxbMCwxLCJDXlxcaW5mdHkoTSkiXSxbMSwxLCJDXntcXGluZnR5fShTKT1cXGZyYWN7Q157XFxpbmZ0eX0oTSl9e0lfe1N9fSJdLFszLDEsIkNee1xcaW5mdHl9KE1fcik9Q15cXGluZnR5KFMpXkciXSxbMSwyLCJcXHN1YnN0YWNre3F1b3RpZW50fSIsMCx7InN0eWxlIjp7ImhlYWQiOnsibmFtZSI6ImVwaSJ9fX1dLFszLDQsIlxcc3Vic3RhY2t7cXVvdGllbnR9IiwwLHsic3R5bGUiOnsiaGVhZCI6eyJuYW1lIjoiZXBpIn19fV0sWzAsMSwiXFxzdWJzdGFja3tyZXN0cmljdGlvbn0iLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJzcXVpZ2dseSJ9fX1dLFs0LDUsIlxcc3Vic3RhY2t7cmVzdHJpY3Rpb259IiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoic3F1aWdnbHkifX19XV0=
\[\begin{tikzcd}
	M & {\mu^{-1}(0)=S} && {M_r} \\
	{C^\infty(M)} & {C^{\infty}(S):=\frac{C^{\infty}(M)}{I_{S}}} && {C^{\infty}(M_r)=C^\infty(S)^G}
	\arrow["{\substack{restriction}}", squiggly, from=1-1, to=1-2]
	\arrow["{\substack{quotient}}", two heads, from=1-2, to=1-4]
	\arrow["{\substack{quotient}}", two heads, from=2-1, to=2-2]
	\arrow["{\substack{restriction}}", squiggly, from=2-2, to=2-4]
\end{tikzcd}\]

where $I_S=\{f\in C^{\infty}(M)~|~f|_S=0\}$.\pause Equivalently:

% https://q.uiver.app/#q=WzAsMyxbMCwwLCJDXlxcaW5mdHkoTSkiXSxbMiwwLCJOX1M9XFx7Zn58fiBmXFxpbiBDXlxcaW5mdHkoTSl8X1N+fkdcXG1hdGhybXstaW52YXJpYW50fVxcfSJdLFs0LDAsIlxcZnJhY3tOX1N9e0lfU30iXSxbMCwxLCJcXHN1YnN0YWNre3Jlc3RyaWN0aW9ufSIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6InNxdWlnZ2x5In19fV0sWzEsMiwiXFxzdWJzdGFja3txdW90aWVudH0iLDAseyJzdHlsZSI6eyJoZWFkIjp7Im5hbWUiOiJlcGkifX19XV0=
\[\small\begin{tikzcd}
	{C^\infty(M)} && {N_S=\{f~|~ f\in C^\infty(M)|_S~~G\mathrm{-invariant}\}} && {\frac{N_S}{I_S}}
	\arrow["{\substack{restriction}}", squiggly, from=1-1, to=1-3]
	\arrow["{\substack{quotient}}", two heads, from=1-3, to=1-5]
\end{tikzcd}\]
\pause

\begin{alertblock}{Problem}
	$C^\infty(S)^G=\frac{N_S}{I_S}$ is not a Poisson algebra in general.
\end{alertblock}
(when constraints not first class, cf. Dirac reduction)
\end{frame}

 \begin{frame}{Other approaches to reduction}

\begin{table}[]
	\begin{tabular}{l|l|l||l}
		algebra $\mathcal{A}_T$ & subalgebra $\mathcal{A}_N$ & ideal $\mathcal{A}_0$ & comment \\
		\hline

		$C^\infty(M)$&      $N_S$     &   $I_S$   &  $\substack{S=\mu^{-1}(0),~ I_S=\{f~|~f|_S=0\}}$ \\
			&&&  $\substack{N_S= \{ f | gf-f\in I_S \forall g\in G\}}$ \\
	& & & \color{red} not Poisson \\ 
	\hline	\pause
	$C^\infty(M)$&   $F_S$         &   $I_S$   &  $\substack{F_S=\{f~|~\{f,I_S\}\subset I_S\}}$ \\
	& & & \color{red} if $I_S\subset F_S$\\
	\hline \pause

	$C^\infty(M)$&   $C^\infty(M)^G$         &   $(I_S)^G$   &  $\substack{C^\infty(M)^G=\{f~|~gf=f\forall g\in G\}}$ \\
	\hline \pause
		$C^\infty(M)$&   $C^\infty(M)^G$         &   $(I_\mu)^G$   & $\substack{I_\mu=\{\sum_{i=1}^k f_i\langle \mu,\xi_i\rangle|~f_i\in C^\infty(M), \xi_i\in \mathfrak g\}}$\\
		\hline \pause
	$C^\infty(M)$&      $N_\mu$     &   $I_\mu$   & $\substack{N_\mu= \{ f | gf-f\in I_\mu \forall g\in G\}}$\\
	\hline

	\end{tabular}\pause
\end{table}
{\bf Remarks }names in order: 'naive' , Dirac, Sniatycki, Arms-Cushman-Gotay, Sniatycki-Weinstein.\pause \\
When $G$ connected $N_\mu=N(I_\mu):=\{f~|~\{f,I_\mu\}\subset I_\mu\}$. 
 \end{frame}

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Algebraic Reduction} 
 a

\end{frame}
\note[itemize]{
	\item  
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Constraint Algebras} 
 a

\end{frame}
\note[itemize]{
	\item  
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%------------------------------------------------------------------------------------------------
\section{Multisymplectic Geometry}
\checkpoint	
%------------------------------------------------------------------------------------------------

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{From symplectic to {multi}symplectic} 
	%
	\begin{center}
		$-$ \emph{multisymplectic means \textbf{going higher} in the degree of $\omega$} $-$
	\end{center}
	\pause
	\begin{defblock}[$n$-plectic manifold ~\emph{(Cantrijn, Ibort, De Le\'on)} \cite{Cantrun2017}]
		\includestandalone[width=0.95\textwidth]{Pictures/Figure_multisym}	
	\end{defblock}
	%
	\pause
		\begin{table}
			\begin{tabular}{c c c}
				symplectic forms \small($n=1$) & $\leftrightsquigarrow$ & volume forms \small($n= \text{dim}(M)-1$)
			\end{tabular}
		\end{table}

	\vfill
	\pause
	\begin{block}{Historical motivation}
		Mechanics: geometrical foundations of \textit{(first-order)} field theories.
	\end{block}
	%
	\begin{table}
		\ifHandout
		%
		\else
		\only<4>{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra &  \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map &  
            \\
            \hline
  \multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &  \multicolumn{1}{@{}c@{}}{}              \\
		\end{tabular}
		}
		\fi
		\onslide<5->{
		\begin{tabular}{|p{0.2\textwidth}|p{0.3\textwidth}|p{0.35\textwidth}|} 
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{mechanics} & \multicolumn{2}{c|}{geometry} \\
            \hline
            \parbox[][20pt][c]{0.2\textwidth}{phase space} & symplectic manifold & multisymplectic manifold \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{classical \\ observables} & Poisson algebra & $L_\infty$-algebra \\[.25em]
            \parbox[][20pt][c]{0.2\textwidth}{symmetries} &  group actions admitting comoment map & group actions admitting (homotopy) comomentum map
            \\
            \hline
  			\multicolumn{1}{c}{}
            &  \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{point-like particles systems}}$} 
            &
            \multicolumn{1}{@{}c@{}}{$\underbrace{\hspace*{.3\textwidth}}_{\text{field-theoretic systems}}$} 
               \\
		\end{tabular}
		}
	\end{table}

\end{frame}
\note[itemize]{
	\item Historically, the interest in multisymplectic manifolds, has been motivated by the need for understanding the geometrical foundations of first-order classical field theories.
	The key point is that, just as one can associate a symplectic manifold to an ordinary classical mechanical system (e.g. a single
point-like particle constrained to some manifold), it is possible to associate a multisymplectic
manifold to any classical field system (e.g. a continuous medium like a filament or a fluid). See frame Extra-\ref{Frame:Ms-Field-Mechanics}

	\item General ideas basic parallelisms with caveats
	\item caveat: points in multiphase spaces are not states
	\item the table hides the duality between geometric and algebraic approaches to the problem.
	\item	Mechanics: geometrical foundations of \textit{(first-order)} field theories.
		\begin{itemize}
		 \item[-] Kijowski, W. Tulczyjew \cite{Kijowski1979}; %(1979)
		 \item[-] Cariñena, Crampin, Ibort \cite{Carinena1991b};% (1991)
		 \item[-] Gotay, Isenberg, Marsden, Montgomery \cite{Gimmsy1};%(1998)
		 \\ $\cdots$
		\end{itemize}
	\item {Def. Multisymplectic manifold $(M,\omega)$}
	$M$ a smooth manifold, $\omega\in \Omega^{n+1}(M)$ closed and non-deg., i.e.:
	\begin{itemize}
		\item $d\omega=0$
		\item $\omega^\flat:TM\to \Lambda^nT^*M, v\mapsto \iota_v\omega$ is injective
	\end{itemize} 
	\item Without the non-deg. condition we call it \emph{premultisymplectic}. 
	\item Examples:
		\begin{itemize} 
			\item orientable $n{+}1$-manifolds with volume, e.g. $S^{n+1}$, $T^{n+1}$, $\mathbb R^{n+1}$ 
			\item multicotangent bunlde $\Lambda^nT^*Q$ of any manifold $Q$, with $\omega=d\theta$. $\theta_\eta(v_1,...,v_n)=\eta(T\pi_Q(v_n),...,T\pi_Q(v_n))$. 
			\item semisimple Lie groups, hyperKähler manifolds,...
		\end{itemize}
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Geometric Multisymplectic Reduction} 
 a

\end{frame}
\note[itemize]{
	\item  
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%------------------------------------------------------------------------------------------------
\section{Multisymplectic observables reduction}
\checkpoint	
%------------------------------------------------------------------------------------------------

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}{Observables in $n$-plectic geometry}
	%
	\begin{defblock}[Hamiltonian $(n-1)$-forms]
		\begin{displaymath}
			\Omega^{n-1}_{ham}(M,\omega) 	:=
			\biggr\{ \sigma \in  \Omega^{n-1}(M) \; \biggr\vert \; 
				\exists \vHam_\sigma \in \mathfrak{X}(M) ~:~ 
				\tikz[baseline,remember picture]{\node[rounded corners,
                        fill=orange!5,draw=orange!30,anchor=base]            
            			(target) {$d \sigma = -\iota_{\vHam_\sigma} \omega$ };
            	}				
				~\biggr\} 
			\end{displaymath}
	\end{defblock}

	\pause
		\tikz[overlay,remember picture]
		{
			\node[rounded corners,
                 fill=orange!5,draw=orange!30,anchor=base]
            	 (base) at ($(current page.north east)-(2,1)$) [rotate=-0,text width=3.5cm,align=center] {\footnotesize{\textcolor{red}{Hamilton-DeDonder-Weyl \\equation}}};
		}	
	\begin{tikzpicture}[overlay,remember picture]
    	\path[->] (base.south) edge[bend right,red](target.north);
    \end{tikzpicture}
	%
	\vfill
	\begin{columns}[T]
		\pause
		\setlength{\belowdisplayskip}{5pt}
		\begin{column}{.50\linewidth}
			%
			\vspace{-1em}
			\begin{thmblock}[Observables $L_\infty$-algebra]
				$\Omega^{n-1}_{ham}(M,\omega)$ endowed with
				\vspace{-.5em}
				\begin{displaymath}
					\lbrace \sigma_1, \sigma_2 \rbrace =			
					~ - \iota_{\vHam_1}\iota_{\vHam_2} \omega 
				\end{displaymath}			
				can be "completed" to a \\ $L_\infty-algebra$.
			\end{thmblock}
			%	
		\end{column}	
		%
		\pause
		\begin{column}{.50\linewidth}
		%
			\begin{itemize}
				\item[\cmark] Skew-symmetric;
				\item[\xmark] multiplication of observables;
				\item[\xmark] Jacobi equation;
				%\\ \hspace*{4.25em} full-fledged Jacobi equation;
				\item[\smark] Jacobi equation \emph{up to homotopies}.
			\end{itemize}			
			%
				\[
					\small
					\{\alpha,\{\beta,\gamma\}\} + \text{\it cyc. perm.}
					= {\color{orange}\d \iota_{X_\alpha} \iota_{X_\beta} \iota_{X_\gamma}\omega}
				\]
		\end{column}	
	\end{columns}
	%
	\vfill
	\pause
	Interesting alternatives:
	\begin{itemize}
		\item  descend to $\Omega^{k-1}_{ham}(M) / {\color{orange}\d\hspace{1pt}\Omega^{k-2}(M)}$, hence getting a Lie algebra;
		\item Incorporate the {\color{orange}discrepancy} as part of the data of the space of observables $ 
				L_\infty(M,\omega) = \Omega_{ham}^{k-1}(M)\oplus{\color{orange}\Omega^{\leq k-2}(M)}$
			%a \textbf{homotopy Lie algebra} or \textbf{$L_\infty$-algebra}.
	\end{itemize}

\end{frame}
\note[itemize]{
	\item
}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\subsection{$L_\infty$-algebra of Observables}
\begin{frame}[fragile,t]{$L_\infty$-algebra of Observables (higher observables) }
	Let be $(M,\omega)$ a $n$-plectic manifold.
	\begin{defblock}[$L_\infty$-algebra of observables ~\emph{(Rogers)} ~\cite{Rogers2010}]

		\hspace{.25em} $L\infty(M,\omega)$ is given by:

		\begin{itemize}
			\item[•] a cochain-complex $(L,\{\cdot\}_1)$ 
		\end{itemize}
		\begin{center}
		\ifHandout
			\includestandalone{Pictures/Figure_Observables}	
		\else
			\includestandalone{Pictures/Frame_Observables}
		\fi				
		\end{center}
		\onslide<2->{
			\begin{itemize}
				\item[•] with $n$ (skew-symmetric) multibrackets $(2 \leq k \leq n+1)$
			\end{itemize}
			\begin{center}
				\includestandalone{Pictures/Equation_Multibracket}	
			\end{center}
		}
		%
	\end{defblock}
  \end{frame}
 \note[itemize]{
	\item if symplectic manifolds are the symmetric take on mechanics, Poisson algebras are the algebraic counterpart.
 	\item A Lie algebra is associated to an ordinary symplectic manifold (its Poisson algebra).
	%(Underlying this is a Lie algebra, whose Lie bracket is the Poisson bracket.)
	Similarly, one associates an Lie-$n$ algebra to any $n$-plectic manifold.
 	% https://ncatlab.org/nlab/show/n-plectic+geometry 	 
 	 %https://ncatlab.org/nlab/show/Poisson+bracket+Lie+n-algebra
	 \item Basically, the higher observables algebra is a chunk of the de Rham complex of $M$ with inverted grading( convention employed here) and an extra structure called "multibrackets".
 	\item ( In the 1-plectic case it reduces to the corresponding Poisson algebra of classical observables)
 	\item Rogers associated to any n-plectic mfd a $L\-\infty$ algebra, Zambon generalized it to the pre-n-plectic case.
 	\item Recognize in the definition of $\{\cdot,\ldots,\cdot\}_k$ the contraction with hamiltonian fields $v_\sigma$ w.r.t. $\sigma$.
  	\item Note $	\iota_{v_{\sigma_1}}\cdots\iota_{v_{\sigma_k}} = (-)^{(k-1)+(k-2)+\dots+1}\iota_{v_{\sigma_k}}\cdots\iota_{v_{\sigma_1}} = (-)^{\frac{k(k-1)}{2}}\iota_{v_{\sigma_k}}\cdots\iota_{v_{\sigma_1}}$ 
 	The definition usually find in literature of Rogers multibrackets involves the coefficient $ (-)^{\frac{k(k-1)}{2}} = -\varsigma(k-1) = (-)^{k+1} \varsigma(k)$.
  \item higher observables is Special instance of a more general object  called $L\-\infty$ Algebra...
 }
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%

%------------------------------------------------------------------------------------------------
% APPENDIX
%------------------------------------------------------------------------------------------------
\appendix
%------------------------------------------------------------------------------------------------

\section{Supplementary Material}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
\begin{frame}
	\begin{center}
	\Huge\emph{Supplementary Material}
	\end{center}
\end{frame}
\note[itemize]{
	\item
}
\addtocounter{framenumber}{-1}
%- . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - . - .%
